---
title: "Data Frames and Functions"
author: "Anna Moeller"
date: "September 19, 2017"
output: 
  html_document:
    theme: cerulean
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data frames
- Each column is a vector
- Find columns by their name with $ (dataframename$columnname)
- subsetting with []

    read [] as *where*, e.g., "find tmp where tmp equals 'sdf' "
    ```{r}
    tmp <- c("a", "sdf", "frog", "sdf")
    tmp[tmp == "sdf"]
    ```
    work from the inside out:
    ```{r}
    tmp == "sdf"
    ```
    so `tmp[tmp == "sdf"]` gives you the value of tmp everywhere `tmp == "sdf" ` is TRUE (e.g., positions 2 and 4)
    
    This also works with other vectors
    ```{r}
    xx <- c(14, 26, 13, 12)
    xx[tmp == "sdf"]
    ```
    gives you the value of xx everywhere `tmp == "sdf"` is TRUE (positions 2 and 4)
    
    This is all important when subsetting data frames
    ```{r}
  # Data.frame
  dat <- data.frame(weight = c(1, 5, 235, pi),
                    name = c("A", "B", "f", "gorilla"),
                    finisher = as.factor(seq(1:4)),
                    date = as.Date(c("2017-04-05", "2017-04-09", "2017-04-01", "2017-03-05")),
                    is.female = c(T, F, T, T),
                    stringsAsFactors = F)
    dat
    ```
    - now there are two dimensions (length and width)
    ```{r}
    dim(dat)
    ```
    every time you subset with [], you must specify both dimensions
    ```{r}
    dat[1, 4]
    dat[1, ]
    dat[, 4]
    dat[dat$name == "gorilla", ]
    ```
    
    - read `dat[dat$word == "shiny", ]` as "the rows of dat *where* the column dat$word is 'shiny', and show me all columns of those rows"
    ```{r}
    dat[, c("name", "is.female")]
    dat[, c(2, 5)]
    ```
    - Using `%in%`
    ```{r}
    dat[dat$name %in% c("gorilla", "f"), ]
    ```

# Functions
- code will always look like
```{r, eval = F}
my.fn <- function(){
  
}
```
- arguments = variables
- default values
- return() (will return only one item from your function)
- Call your function with
```{r, eval = F}
newvar <- my.fn()
```
and newvar will take on the value of the thing your function returns 

# Practice
1. Brainstorm at least 5 words that describe what you want to learn about R this semester
2. Create a data frame called dat, following this example:

```{r echo = F}
  dat <- data.frame(name = "Anna",
                    word = c("shiny", 
                             "web scraping", 
                             "automating",
                             "tidyverse", 
                             "Git"), 
                    order = 1:5,
                    stringsAsFactors = F)
  knitr::kable(dat, 
               caption = "Example Data Frame", 
               padding = 0)
               # align = "l", 
               # format = "markdown", 
               # )

  # ```{r , results = "asis" }
  # xtable::xtable(dat, type = "html", caption = "xtable")
  # print(xtable::xtable(dat, caption = "Example 2"),
  #       type = "html",
  #       html.table.attributes = "border=0")
  # stargazer::stargazer(dat, summary = F, rownames = F, align = T,
  #                      type = "html", title = "Example data frame")
```
```{r, echo = F}
 my.fn <- function(name){
   out <- data.frame(name = name,
                    word = c("shiny", 
                             "web scraping", 
                             "automating",
                             "tidyverse", 
                             "Git"), 
                    order = 1:5,
                    stringsAsFactors = F)
   return(out)
 }
dat2 <- my.fn(name = "AnnA")

```

3. Save this as an RData file. Practice closing your R session and loading dat without re-running your code. 

### Notes
1. Make sure that your column names match mine exactly
2. Make sure that your columns are characters, not factors
3. Make everything lower case, unless it is a proper noun

### Further practice (use Google, help files, and your neighbors!)
- See what happens if your column names are not all the same length, when defining your data frame
- Practice changing an entire column at once from lower case to upper case
- Send your RData file to a friend and combine the two data.frames. 
- Change one person's columns to factors and see what happens if you try to combine the two data frames
- Subset 
    - Subset your data to only the rows where order = 3 or 4
    - Subset your data to only columns 1 and 3
- Change your data.frame to a tibble. Compare and contrast the two
- Turn this code into a function that takes no arguments. Turn it into a function that takes the argument "name" and fills in the first column of your data frame. Practice calling your function. Assign the output of your function to a new object named "dat2"

### More advanced practice
- Try changing the order of your rows using the function order()
- Use gsub() to change one of your column names
- Create a word cloud from your data frame
